---
title: 'Gapminder Data ranging: Homework 2'
date: '2019-09-22'
output: html_document
---
# Setting Up
Loading needed Packages
```{r, warning= FALSE, message= FALSE}
library(gapminder)
library (tidyverse)
```


# Exercise 1: 1.1
Subsetting the gapminder dataset to three countries of choice
- Canada, United Kingdom and Germany - during the 1970s.
 
```{r}
filter(gapminder, 
       country == 'Canada'| country== 'United Kingdom' | country== 'Germany') %>%
  filter(year %in% (1970:1979)) %>%
  knitr::kable()
```

# Exercise 1: 1.2

Selecting the GDP per Capita and Population from the subset of the
gapminder data created in Exercise 1: 1.1
```{r}
filter(gapminder, country == 'Canada'| country == 'United Kingdom'| country== 'Germany') %>% 
  filter(year %in% (1970:1979)) %>%
  select(country,gdpPercap) %>%
  knitr::kable()
```

# Exercise 1: 1.3

Subset the gapminder dataset by countries that have experienced a drop in their
Life Expectancy
```{r}
gapminder %>%
  arrange(year) %>%
  group_by(country) %>%
   mutate (increaseLifeExp =  lifeExp-lag(lifeExp)) %>% 
   filter( (increaseLifeExp<0)) %>%
   DT::datatable() 

```

# Exercise 1: 1.4

Subset of gapminder dataset with max GDP per capita for each country
```{r}
gapminder %>% 
  group_by(country) %>%
  filter(gdpPercap== max(gdpPercap)) %>%
  DT::datatable() 
```

# Exercise 1 : 1.5

Plotting Life Expectancy  vs GDP per Capita in Canada
```{r}
gapminder %>%
  select (country,gdpPercap,lifeExp) %>%
  filter (country=='Canada') %>%
  ggplot(aes(lifeExp,gdpPercap)) +  geom_point(alpha = 0.5) +
  scale_x_continuous("Life Expectancy") +
  theme_bw() + scale_y_log10("GDP per capita", labels= scales::dollar_format()) + 
ggtitle("Life Expectancy vs GDP per capita in Canada")
```

# Exercise 2:

Continuous variables= gdpPercap

Discrete Variable= continent

### Possible Values for continent:
```{r}
summary(gapminder$continent)
```

### Distribution of Continents:
Number of observations for each continent in the gapminder dataset
```{r}
gapminder %>%
  ggplot( aes(continent)) + geom_bar() +ylab("Count")+xlab("Continent")+ggtitle("Distribution of Observations in Gapminder Data Set by Continent")
```


### Possible Values for gdpPercap
```{r}
summary(gapminder$gdpPercap)  
```
### Mean Value for gdpPercap
```{r}
mean(gapminder$gdpPercap)  
```

### Standard Deviation of gdpPercap
```{r}
sd(gapminder$gdpPercap)
```

### Spread of GDP per capita values by continent:
The spread of GDP per cpita values for each continent in the Gapminder dataset
```{r}
boxplot(gdpPercap~continent,data=gapminder, main="Boxplots of Gapminder GDP per Capita by Continent", 
    xlab= "Continent", ylab="GDP per Capita")
```

### Distribution of GDP Per Capita

Histogram of the distribution of gdpPercap in the gapminder dataset
```{r, message= FALSE, warning= FALSE}
gapminder %>%
  drop_na() %>%
ggplot() + geom_histogram( mapping=aes(gdpPercap)) + ylab("Count") + 
  xlab("GDP per Capita") + ggtitle("Histogram of GDP per Capita Value Distribution")
```

# Exercise 3: Plotting the Air Quality Dataset

The airquality dataset can be loaded from the package datasets using the following code:

```{r, results="hide"}
datasets::airquality
```

### Scatter plot of  Mean Ozone levels vs Temperature(°C) in airquality dataset
```{r}
airquality %>%
  filter(!is.na(Month) & !is.na(Ozone))%>%
ggplot(aes(Ozone,Temp)) + geom_point(alpha=0.5) +  scale_x_continuous("Ozone (ppb)") +
theme_bw() + ylab("Temperature (°C)") + ggtitle("Mean Ozone (ppb) vs Temperature(°C)")

```

### Histogram of Air quality data set
```{r, warning=FALSE, message=FALSE}
airquality %>%
  drop_na() %>%
  ggplot() + aes(Ozone)+ geom_histogram()+ggtitle("")
```


### Bonus
The code described in the assignment does not achieve the desired result.
It returns only half the entries needed. 
It skips every second entry for Rwanda and every first for Afghanistan.
This is because the filter function was looking for Rwanda and Afghanistan together and not first for all the entries of Rwanda and then all the entries of Afghanistan
That is, it recycles the c("Rwanda","Afghanistan") vector. 
It looks for Rwanda, then Afghanistan, then Rwanda and so on.Rather than for Rwanda or Afghanistan.

Corrected Version:

```{r}
gapminder %>%
  filter( country == "Rwanda"| country== "Afghanistan") %>%
  DT::datatable()
```


