---
title: "Gapminder Data Set Exploration"
author: "Denitsa Vasileva"
date: '2019-09-14'
output: github_document
---

# Loading the gapminder data set
gapminder is a built-in data set in R. To load it we use the following code:

```{r}
library(gapminder)
```

# Exploring the gapminder data set
## Summary statistics
The following code provides summary statistics for the gapminder data set.
```{r}
summary(gapminder)
```
## How many countries are represented in this data set?
To find how many countries are represented in the data set we use the following code:
```{r}
num_countries<- length(unique(gapminder$country))
print (num_countries)
```
This shows that 142 countries were included in this study.

## What was the country with the lowest GDP per capita overall?

```{r}
min_GDP_country <- data.frame(gapminder[gapminder$gdpPercap==min (gapminder$gdpPercap),])
print (min_GDP_country )
```
## What was the country with the highest GDP per capita?

```{r}
max_GDP_country <- data.frame(gapminder[gapminder$gdpPercap==max (gapminder$gdpPercap),])
print (max_GDP_country)
```
## What was the country with the lowest GDP per capita for each of the years the information was recorded?
```{r}
df_min_GDP_year <- data.frame()
years<- unique(gapminder$year)
for (year in years){
  new_df<-data.frame(gapminder[gapminder$year==year,])
  x<-new_df[(new_df$gdpPercap==min(new_df$gdpPercap)),]
  df_min_GDP_year<-rbind(df_min_GDP_year,x )
}
print(df_min_GDP_year)
```
## What was the country with the highest GDP per capita for each of the years the information was recorded?
```{r}
df_max_GDP_year <- data.frame()
years<- unique(gapminder$year)
for (year in years){
  new_df<-data.frame(gapminder[gapminder$year==year,])
  x<-new_df[(new_df$gdpPercap==max(new_df$gdpPercap)),]
  df_max_GDP_year<-rbind(df_max_GDP_year,x )
}
print(df_max_GDP_year)
```

