---
title: "Gapminder Data Set Exploration"
author: "Denitsa Vasileva"
date: '2019-09-14'
output:
  beamer_presentation: default
  ioslides_presentation: default
---

## What is the gapminder data set?
* Built-in data set in R
*  A Tibble
*  Contains GDP of countries over a 12 year period
* To load it we use the following code:

```{r, fig.align='right', fig.height=4}
library(gapminder)
```


## Exploring the gapminder data set: Summary Statistics
To get summary statistics on the gapmind data set run the following code:
```{r, results= "hide"}
   summary(gapminder)
```
## Exploring the gapminder data set: How many countries are included in this data set?
To find how many countries are represented in the data set we use the following code:
```{r}
num_countries <- length(unique(gapminder$country))
print (num_countries)
```
142 countries were included in this study.

## Exploring the gapminder data set: What was the country with the lowest GDP per capita overall?

```{r,results="hide"}
min_GDP_country <- data.frame(gapminder
[gapminder$gdpPercap==min (gapminder$gdpPercap),])
print (min_GDP_country )
```
 The Dem. Republic of Congo was the country with the lowest GDP per capita in this data set.
 
## Exploring the gapminder data set: What was the country with the highest GDP per capita?

```{r}
max_GDP_country <- data.frame(gapminder
[gapminder$gdpPercap==max (gapminder$gdpPercap),])
print (max_GDP_country)
```
Kuwait was the country with the highest GDP per capita in this data set.

## Exploring the gapminder data set: Minimum GDP per capita per year

To find the lowest GDP per capita in every year use the following code:

```{r,results="hide" }
df_min_GDP_year <- data.frame()
years <- unique(gapminder$year)
for (year in years){
  new_df<-data.frame(gapminder
                     [gapminder$year==year,])
  x<-new_df[(new_df$gdpPercap==
               min(new_df$gdpPercap)),]
  df_min_GDP_year<-rbind(df_min_GDP_year,x )
}
print(df_min_GDP_year)
```


## Exploring the gapminder data set: Maximum GDP per capita per year

To find the highest GDP per capita in every year use the following code:

```{r, results="hide"}
df_max_GDP_year <- data.frame()
years<- unique(gapminder$year)
for (year in years){
  new_df<-data.frame(gapminder
                     [gapminder$year==year,])
  x<-new_df[(new_df$gdpPercap==max
             (new_df$gdpPercap)),]
  df_max_GDP_year<-rbind(df_max_GDP_year,x )
}
print(df_max_GDP_year)
```
