---
title: 'Homework 3: Gapminder exploration with dplyr and ggplot'
date: '2019-09-29'
output:
  html_document:
    toc: true

---

```{r setup, result="hide", message=FALSE}
library(tidyverse)
library(gapminder)
library(ggplot2)
library(DT)
```


# Task #2
<style type="text/css">
.twoC {width: 100%}
.clearer {clear: both}
.twoC .table {max-width: 50%; float: right}
.twoC img {max-width: 50%; float: left}
</style>



<div class="twoC">
```{r, echo=FALSE}
minmaxGDP<-gapminder %>%
  group_by(continent) %>%
  summarize(max_GDP=max(gdpPercap),min_GDP=min(gdpPercap))
knitr::kable(minmaxGDP)
```



```{r,echo=FALSE}
gapminder %>%
  ggplot(aes(continent,gdpPercap,colour=continent))+geom_boxplot(alpha=0.5)+ggtitle("Boxplot of GDP per Capita per Continent") + xlab("Continent")+ ylab("GDP Per Capita")+ scale_y_continuous(labels = scales::dollar_format())+scale_color_discrete("Continent")
```
</div>
<div class="clearer"></div>
## Explanation

The tibble is a tabular representation of the minimum and maximum GDP per Capita for each continent in the gapminder dataset.
The boxplot shows the minimum and maximum but also provides additional useful information such as  median GDP per capita for each continent as well as the 1st quartile (25th percentile) and 3rd quartile (75th percentile) for GDP per capita for each continent. The boxplot also clearly shows outliers which we cannot see from measure such as maximum and minimum.
As can be seen from the plot, Oceania has the highest median GDP per capita while Africa has the lowest. Asia has the greatest spread of GDP per capita values while Oceania has the least. 

# Task #3

## Tibble of the Mean and Standard Deviation per Continent

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(Mean=mean(gdpPercap), StandardDev=sd(gdpPercap))%>%
  knitr::kable()
```

## Histogram of Spread of the GDP Per Capita Per Continent
```{r}
ggplot(gapminder, aes(gdpPercap))+facet_wrap(~ continent, nrow=4, scales="free_y") +
  geom_histogram(bins = 50)+ggtitle("Histograms of GDP Per Capita Per Continent ") + xlab("GDP Per Capita")+scale_x_continuous(labels=scales::dollar_format())+ylab("Frequency")
```

## Density Diagram of the GDP Per Capita Per Continent

```{r}
ggplot(gapminder, aes(gdpPercap))+facet_wrap(~ continent, nrow=4, scales="free_y") +
  geom_density()+ggtitle("Density Diagram of GDP Per Capita Per Continent ") + xlab("GDP Per Capita")+scale_x_continuous(labels=scales::dollar_format())+ylab("Frequency")
```

## Explanation

The tibble shows the mean and the Standard Deviation of the GDP per Capita for each continent.
The histograms show the spread of the GDP per Capita values for each continent. From the histograms we can see that Asia has the largest spread of GDP per capita values while Africa has the smallest spread. In addition, we can see that the values for Africa are left-skewed. Asia has a significant right-tail as do the Americas meaning that their distribution is skewed to the right.
The density diagram complements the histogram by better demonstrating the shape of distribution of the data. For example, density plots allow to more easily determine whether the data follows a normal distribution or is skewed.

# Task #4
## Tibble of Weighted (by population) Mean Life Expectancy per year
```{r}
min_life_exp<-gapminder %>%
  group_by(year)%>%
  summarise( 
            lifeExp_weighted_mean = weighted.mean(lifeExp,pop)) 
  knitr::kable(min_life_exp)
```

## Bar Graph of Weighted (by population) Mean Life Expectancy per year

```{r}
min_life_exp %>%
ggplot(aes(x=year,y=lifeExp_weighted_mean))+geom_col()+xlab("Year")+ylab("Weighted Mean Life Expectancy")+ ggtitle("Weighted Mean Life Expetancy Per Year")

```

## Explanation
The tibble shows the mean life expectancy per year when weighted by population. The bar graph shows that the trend is towards an increasein the mean weighted life expectancy over the years of the gapminder data set.
